cardinalR: Generating interesting high-dimensional data structures

### 2025-04-30 paper-cardinalR.Rmd
Added the colors for NLDR layouts

```{r layouts}
clusters <- as.vector(five_clusts[, 5])

tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")

nldr1 <- tsne_data |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")

nldr2 <- umap_data |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("b", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")

nldr3 <- phate_data |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("c", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")

nldr4 <- trimap_data |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("d", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")

nldr5 <- pacmap_data |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("e", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")

nldr6 <- pca_data |>
  ggplot(aes(x = pca1,
             y = pca2))+
  geom_point(alpha=0.1, size=1, colour='#000000') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1) 
```

```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}
nldr1 + nldr2 + nldr3 +
  nldr4 + nldr5 + nldr6 +
  plot_layout(ncol = 3)
```

>>>>

```{r layouts}
clusters <- five_clusts |> pull(cluster)

tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds") |>
  mutate(cluster = clusters)

nldr1 <- tsne_data |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("a", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10))

umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds") |>
  mutate(cluster = clusters)

nldr2 <- umap_data |>
  ggplot(aes(x = UMAP1,
             y = UMAP2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("b", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10))

phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds") |>
  mutate(cluster = clusters)

nldr3 <- phate_data |>
  ggplot(aes(x = PHATE1,
             y = PHATE2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("c", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10))

trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds") |>
  mutate(cluster = clusters)

nldr4 <- trimap_data |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("d", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10))

pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds") |>
  mutate(cluster = clusters)

nldr5 <- pacmap_data |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("e", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10))

pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds") |>
  mutate(cluster = clusters)

nldr6 <- pca_data |>
  ggplot(aes(x = pca1,
             y = pca2,
             color = cluster))+
  geom_point(alpha=0.1, size=1) +
  interior_annotation("f", c(0.08, 0.93)) + 
  scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3','#e7298a','#66a61e')) +
  theme(aspect.ratio = 1,
        legend.text = element_text(size = 10)) 
```

```{r, label = "fig-nldr-layouts", fig.cap="Six different dimension reduction representations of the `mobiusgau` data using default hyperparameter settings: (a) tSNE, (b) UMAP, (c) PAHTE, (d) TriMAP, (e) PaCMAP, and (f) PCA."}

nldr1 + nldr2 + nldr3 +
  nldr4 + nldr5 + nldr6 +
  plot_layout(ncol = 3, guides = "collect")
```
